ABM simulating short-term mating behaviors

### 2022-06-10 submission/manuscript.tex
final edits

\hypertarget{an-agent-based-model-of-gender-differences-in-short-term-mating-behaviors-as-a-result-of-mating-preferences}{%
\section*{An Agent-Based Model of Gender Differences in Short-Term
Mating Behaviors as a Result of Mating
Preferences}\label{an-agent-based-model-of-gender-differences-in-short-term-mating-behaviors-as-a-result-of-mating-preferences}}
\addcontentsline{toc}{section}{An Agent-Based Model of Gender
Differences in Short-Term Mating Behaviors as a Result of Mating
Preferences}

\vspace{30mm}

Yurun Ying\textsuperscript{1,\(\dagger\)}, Jan
Antfolk\textsuperscript{2}, Pekka Santtila\textsuperscript{1,3}

\textsuperscript{1} Faculty of Arts and Sciences, NYU Shanghai

\textsuperscript{2} Faculty of Arts, Psychology and Theology, Åbo
Akademi University

\textsuperscript{3} NYU-ECNU Institute for Social Development at NYU
Shanghai

@@

\vspace{40mm}

\textsuperscript{\(\dagger\)} To whom correspondence should be
addressed: \url{yurun.ying@nyu.edu}

@@

Gender differences in short-term mating behaviors are well-documented in
human sexuality research. Existing studies usually conflate gender
differences in mating preferences with differences in sexual behaviors,
which is theoretically problematic. Using an agent-based model, we
investigated the circumstances under which heterosexual and homosexual
men and women's differential preferences for short-term mating would
result in gender differences in short-term mating behaviors. The model
showed that when all individuals in a closed heterosexual population
were considered, men and women had the same average number of short-term
mating experiences and short-term mates even when men had stronger
preferences for short-term mating. Men (vs.~women) had a higher average
number of both experiences and mates only when heterosexual men and
women who successfully participated in the mating pool (i.e., those with
a non-zero number of short-term mating experiences) were considered.
Moreover, when men (vs.~women) had stronger preferences for short-term
mating, gay men had a higher average number of both experiences and
mates compared to both lesbian women and heterosexual men. These results
suggest that theoretically speaking, even when gender differences in
mating preferences exist, those in short-term mating behaviors only
occur among particular populations, or when men's preferences for
short-term mating are not constrained by those of women. Suggestions for
future research in human mating psychology and behaviors were provided.

@@

\emph{Keywords:} agent-based modeling, short-term mating, casual sex,
sexual strategy theory, female choice hypothesis

@@

As a starting point, we took gender differences in mating preferences as
conceptualized by the sexual strategy theory (Buss \& Schmitt, 1993;
Buss \& Schmitt, 2019), which is repeatedly supported by empirical
investigations (e.g., Schmitt, 2003; Walter et al., 2020). Using an
agent-based model, we investigated whether men and women's differential
preferences for short-term mating would result in gender differences in
short-term mating behaviors, and if so, then under what conditions they
would result in differences in the number of short-term mating
experiences and short-term mates.

@@

Gender differences in mating preferences have been studied in the light
of sexual strategy theory (Buss \& Schmitt, 1993; Buss \& Schmitt,
2019). It posits that since the minimum obligatory investment that men
must devote to their offspring (contribution of sperm through one sexual
act) is lower than that of women (gestation, labor, and lactation), men
tend to be relatively more interested in short-term mating compared to
women. This is because the number of offspring women can produce is
limited and cannot be increased by mating with a large number of men,
whereas the reverse is the case for men. As a result, men are predicted
to not only have a greater interest in short-term mating, but also
desire a larger number of short-term mates in a given period and be less
selective with respect to accepting a potential mate (Buss \& Schmitt,
1993).

The hypotheses derived from sexual strategy theory have received
extensive support in empirical studies. For example, men have reported
to be currently seeking short-term mates to a larger extent than women
(Buss \& Schmitt, 1993; Schmitt et al., 2001; Schmitt, 2003), and a
larger proportion of men were in any way seeking short-term mates
(vs.~not seeking) (Schmitt, 2003). Similar results have been found among
both U.S. college students (Buss \& Schmitt, 1993; Schmitt et al., 2001)
as well as in ten world regions in a cross-cultural study (Schmitt,
2003). Similarly, men across the world have also reported to desire more
short-term sexual partners compared to women over different future time
periods (e.g., a month, a year) (Buss \& Schmitt, 1993; McBurney et al.,
2005; Schmitt, 2003). The gender differences have been found regardless
of whether they were estimated using mean (Buss \& Schmitt, 1993;
Schmitt et al., 2001; Schmitt, 2003), median (Schmitt et al., 2001;
Schmitt, 2003), or percentage (Schmitt, 2003) statistics.

As for mating standards, men are less selective in accepting someone as
a potential short-term mate. For example, studies using U.S. college
samples have found that the minimum percentile ranks of the desirability
of a potential short-term sexual partner accepted by men were lower than
those accepted by women, in terms of both the overall desirability and
the desirability of individual traits (e.g., social status,
attractiveness) (Kenrick et al., 1990; Regan, 1998). Another study also
found that when presented with identical descriptions of potential
mates, men on average rated them as more desirable than women did
(Wiederman \& Dubois, 1998).

@@

Since heterosexual sex in most cases involves one man and one woman,
men's relatively high interest in short-term mating can be constrained
by women's preference for long-term relationships (Archer, 2019; Symons,
1979). This is because men's short-term mating preferences -- with the
exception of sexual violence -- can only translate into behaviors when
there are women willing to have sex with them. When a new short-term
mating encounter occurs, it counts towards both men's total number of
short-term mating encounters as well as women's, and both men and women
have new short-term mates. Therefore, the total number of short-term
mating experiences and short-term mates must be equal between
heterosexual men and women at the population level. In a closed
population with equal sex ratio, it follows that the average number of
experiences and mates should also be equal between men and women.
Following this line of reasoning, we would expect to observe no gender
difference in short-term mating behaviors among heterosexual individuals
even when there were gender differences in mating preferences.

However, it is important to note that although heterosexual men's total
number of short-term mating is equal to that of women, this does not
necessarily mean that the total number of men and women who have ever
had short-term mating must be equal. Empirical studies have found that
the proportion of individuals who are sexual inexperienced is slightly
higher among men than women in most age groups above 18 years old
(Ghaznavi et al., 2019; Haydon et al., 2014). As a postulation, there
may be a smaller proportion of men (vs.~the proportion of women) who
participate in the mating pool and contribute to the total number of
short-term mating. For example, suppose that there is a population of 20
individuals with an equal sex ratio. Heterosexual short-term mating has
happened \(n\) times in this population, but only 3 men and 8 women have
had any short-term mating experiences. Since the total number of
short-term mating behaviors is the same between men and women, we would
expect that in a population of individuals who had any short-term
mating, the average number of short-term mating experiences and mates
was higher among men than women (e.g., in the aforementioned example,
the average number of short-term experiences is \(\frac{n}{3}\) for men
and \(\frac{n}{8}\) for women who are in the mating pool).

As a comparison, in the cases of gay men and lesbian women, men's
preferences are not constrained by women's, but only by those of other
men, who, arguably, have a similar high interest in short-term mating.
This would allow for a more direct behavioral expression of men's mating
preferences. The notion that gay men have less restricted preferences is
borne out in the proportion of individuals who have engaged in
extradyadic sex (Peplau et al., 2004). A study showed that the
proportions of heterosexual men and women who had engaged in extradyadic
sex were 26\% and 21\%, respectively, while the statistics among gay men
and lesbian women were 82\% and 28\%, respectively (Peplau et al.,
2004). Therefore, we would expect to observe gender differences in
short-term mating behaviors between gay men and lesbian women. Moreover,
we would also expect to find that gay men, compared to heterosexual men,
engaged in more short-term mating behaviors due to less constraints on
their mating preferences.

@@

We also modeled forming long-term relationships as a background process.
A pair of individuals had a possibility to commit to a long-term
relationship upon meeting each other. This process was modeled at a
behavioral level and the psychological processes behind it was ignored,
although we recognized gender differences may also exist in long-term
mating strategy (Buss \& Schmitt, 1993; Buss \& Schmitt, 2019).

We modeled these processes among both heterosexual individuals and gay
men and lesbian women to examine whether constraints set by the opposite
sex's preferences would affect short-term mating behaviors. Individuals'
sexual orientation was conceptualized in terms of behaviors only in our
model. Heterosexual men had short-term mating or formed a long-term
relationship only with women, while gay men only with other men, and
mutatis mutandis for women.

We formulated the following hypotheses in the present study: assuming
gender differences in preferences for short-term mating, 1) there would
be no gender differences in short-term mating behaviors between
heterosexual men and women when the full population was considered, 2) a
smaller proportion of heterosexual men (vs.~heterosexual women) had ever
engaged in short-term mating, 3) in a population of heterosexual
individuals who had any short-term mating, men had higher average number
of short-term mating behaviors than women, 4) gay men would engage in
more short-term mating behaviors compared to lesbian women, and 5) gay
men would engage in more short-term mating behaviors compared to
heterosexual men.

@@

The model measured two outcomes: (1) the number of short-term mating
experiences of men and women; (2) the number of short-term mates of men
and women. Additionally, we also measured the number of men and women in
the mating pool (i.e., those with a non-zero number of short-term mating
experiences). The average numbers of short-term mating experiences and
short-term mates were calculated by taking the average among the whole
population of men and women (\emph{N\textsubscript{men}} =
\emph{N\textsubscript{women}} = 150) and among those who had engaged in
any short-term mating.

@@

At each time step, agents first checked whether they were in a long-term
relationship. If they were not in a long-term relationship, they set
their heading randomly (if they were within the movement range from the
birthplace) or faced the birthplace (if they were out of the movement
range from the birthplace) and moved by a random distance. The random
distance was less than half of the movement range. If they were in a
long-term relationship, they did not move. Then, the agents decided on
whether to engage in short-term mating using the predetermined
likelihood (either 25\% or 40\%).

@@

Then, women randomly selected one man at the same location as their
potential long-term partner. There was a 10\% of chance that a pair
would form a long-term relationship. After forming a long-term
relationship, both men and women changed to coupled status and
registered each other as their long-term partner.

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=0.8\columnwidth]{figures/fig2_men_vs_women_inpool.pdf}
  \caption{\textbf{Short-term mating behaviors of heterosexual men and women in the mating pool after 1,000 time steps in the model when gender differences existed in mating preferences.} Violin plots summarizing the outcome variables. Plot (A) shows the average number of short-term mating experiences in the mating pool, and plot (B) shows the average number of short-term mates in the mating pool. Central points show mean values and whiskers represent standard errors, but the standard errors are small and overlap to form a single bar. The averages were calculated using the population of heterosexual men and women in the mating pool (proportion: \textit{M\textsubscript{men}} = 0.38, \textit{M\textsubscript{women}} = 0.51).}
  \label{fig:men_vs_women_inpool}
\end{figure}

@@

Gay men engaged in short-term mating behaviors more than heterosexual
men did (Figure \ref{fig:hetero_vs_gay}). The average number of
short-term mating experiences (as calculated using the full population
of men) was higher among gay men (\emph{M} = 2.63, \emph{SD} = 0.37)
than among heterosexual men (\emph{M} = 1.43, \emph{SD} = 0.24),
\emph{t}(4,998) = --134.87, \emph{p} \textless{} .001, Cohen's \emph{d}
= 3.81, 95\% CI = --1.21 to --1.18. Similarly, the average number of
short-term mates (as calculated using the full population of men) was
higher among gay men (\emph{M} = 1.72, \emph{SD} = 0.19) than among
heterosexual men (\emph{M} = 0.80, \emph{SD} = 0.11), \emph{t}(4,998) =
--206.91, \emph{p} \textless{} .001, Cohen's \emph{d} = 5.85, 95\% CI =
--0.92 to --0.91.

\begin{figure}[h]
  \centering
  \includegraphics[width=0.8\columnwidth]{figures/fig3_hetero_vs_gay_men.pdf}
  \caption{\textbf{Short-term mating behaviors of heterosexual and gay men after 1,000 time steps in the model when gender differences existed in mating preferences.} Violin plots summarizing the outcome variables. Plot (A) shows the average number of short-term mating experiences, and plot (B) shows the average number of short-term mates. Central points show mean values and whiskers represent standard errors, but the standard errors are small and overlap to form a single bar. The averages were calculated using the full population of heterosexual and gay men in the model (\textit{N\textsubscript{hetero}} = \textit{N\textsubscript{gay}} = 150).}
  \label{fig:hetero_vs_gay}
\end{figure}

\hypertarget{comparing-across-conditions}{%
\subsection{Comparing across
conditions}\label{comparing-across-conditions}}

@@

\caption{\label{tab:table_1}\textbf{Short-term mating behaviors of heterosexual men and women in the mating pool after 1,000 time steps in the model}}
\centering
\begin{threeparttable}
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{2in}llllllllll}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{5}{c}{\makecell[c]{Average short-term mating\\experiences (in-pool)}} & \multicolumn{5}{c}{\makecell[c]{Average short-term\\mates (in-pool)}} \\
\cmidrule(l{3pt}r{3pt}){2-6} \cmidrule(l{3pt}r{3pt}){7-11}
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Men} & \multicolumn{2}{c}{Women} & \multicolumn{1}{c}{ } & \multicolumn{2}{c}{Men} & \multicolumn{2}{c}{Women} & \multicolumn{1}{c}{ } \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){7-8} \cmidrule(l{3pt}r{3pt}){9-10}
\multicolumn{1}{c}{\em{ }} & \multicolumn{1}{c}{\em{M}} & \multicolumn{1}{c}{\em{SD}} & \multicolumn{1}{c}{\em{M}} & \multicolumn{1}{c}{\em{SD}} & \multicolumn{1}{c}{\em{d}} & \multicolumn{1}{c}{\em{M}} & \multicolumn{1}{c}{\em{SD}} & \multicolumn{1}{c}{\em{M}} & \multicolumn{1}{c}{\em{SD}} & \multicolumn{1}{c}{\em{d}}\\
\midrule
Same likelihood*Same standard & 2.21 & 0.38 & 2.20 & 0.37 & 0.03 & 1.39 & 0.12 & 1.39 & 0.12 & 0.04\\
Same likelihood*Different standard & 2.82 & 0.36 & 2.21 & 0.28 & 1.89 & 1.76 & 0.15 & 1.38 & 0.09 & 3.13\\
Different likelihood*Same standard & 2.80 & 0.48 & 2.78 & 0.47 & 0.03 & 1.57 & 0.14 & 1.56 & 0.14 & 0.05\\
Different likelihood*Different standard & 3.78 & 0.50 & 2.79 & 0.35 & 2.31 & 2.12 & 0.17 & 1.56 & 0.11 & 3.84\\
\bottomrule
\end{tabular}}
\begin{tablenotes}
\small
\item \textit{Note.} The averages were calculated with subsamples of men and women who had had engaged in short-term mating behaviors. See Supplementary Materials for the mean and standard deviation of sample sizes.
\end{tablenotes}
\end{threeparttable}
\end{table}

Among gay men and lesbian women, gender differences in short-term mating
behaviors emerged both when they had different short-term mating
likelihood and when their mating standards were different. Gay men
(vs.~lesbian women) had a higher average number of short-term mating
experiences and short-term mates when they had a higher short-term
mating likelihood (experiences: \emph{M\textsubscript{men}} = 0.80,
\emph{SD\textsubscript{men}} = 0.23, \emph{M\textsubscript{women}} =
0.32, \emph{SD\textsubscript{women}} = 0.10, Cohen's \emph{d} = 2.67;
partners: \emph{M\textsubscript{men}} = 0.52,
\emph{SD\textsubscript{men}} = 0.13, \emph{M\textsubscript{women}} =
0.25, \emph{SD\textsubscript{women}} = 0.07, Cohen's \emph{d} = 2.60),
or when they had lower mating standards (experiences:
\emph{M\textsubscript{men}} = 1.05, \emph{SD\textsubscript{men}} = 0.18,
\emph{M\textsubscript{women}} = 0.32, \emph{SD\textsubscript{women}} =
0.10, Cohen's \emph{d} = 5.10; partners: \emph{M\textsubscript{men}} =
0.82, \emph{SD\textsubscript{men}} = 0.12, \emph{M\textsubscript{women}}
= 0.25, \emph{SD\textsubscript{women}} = 0.07, Cohen's \emph{d} = 5.78).
In comparison, virtually no gender differences in short-term mating
behaviors existed when gay men and lesbian women had the same short-term
mating likelihood and mating standards (Table 2).

@@

The present study used agent-based modeling to investigate whether men
and women's differential preferences for short-term mating would result
in gender differences in short-term mating behaviors, and if so, then
under what conditions they would result in differences in the number of
short-term mating experiences and short-term mates. We formulated five
hypotheses: 1) there would be no gender differences in short-term mating
behaviors between heterosexual men and women when the full population
was considered, 2) a smaller proportion of heterosexual men
(vs.~heterosexual women) had ever engaged in short-term mating, 3) in a
population of heterosexual individuals who had any short-term mating,
men had higher average number of short-term mating behaviors than women,
4) gay men would engage in more short-term mating behaviors compared to
lesbian women, and 5) gay men would engage in more short-term mating
behaviors compared to heterosexual men. The results from 1,000 time
steps in a model simulating men and women's mating behaviors provided
strong evidence in favor of our hypotheses. First of all, we found no
gender differences in short-term mating behaviors between heterosexual
men and women when the full population was considered. However, compared
to heterosexual women, there was a smaller proportion of heterosexual
men participating in the mating pool, but they had higher average
numbers of short-term mating experiences and short-term mates. Secondly,
as compared to lesbian women, gay men had higher average numbers of
short-term mating experiences and short-term mates. Finally, gay men
also had higher average numbers of short-term mating experiences and
short-term mates as compared to heterosexual men.

>>>>

\hypertarget{an-agent-based-model-of-gender-and-sexual-orientation-differences-in-short-term-mating-behaviors-as-a-result-of-mating-preferences}{%
\section*{An Agent-Based Model of Gender and Sexual Orientation
Differences in Short-Term Mating Behaviors as a Result of Mating
Preferences}\label{an-agent-based-model-of-gender-and-sexual-orientation-differences-in-short-term-mating-behaviors-as-a-result-of-mating-preferences}}
\addcontentsline{toc}{section}{An Agent-Based Model of Gender and Sexual
Orientation Differences in Short-Term Mating Behaviors as a Result of
Mating Preferences}

@@

@@

Gender differences in short-term mating behaviors are well-documented in
human sexuality research. Existing studies usually conflate gender
differences in mating preferences with differences in sexual behaviors,
which is theoretically problematic. Using an agent-based model, we
investigated the circumstances under which men and women's differential
preferences for short-term mating would result in gender differences in
short-term mating behaviors. The model showed that when all individuals
in a closed heterosexual population were considered, men and women had
the same average number of short-term mating experiences and short-term
mates even when men had stronger preferences for short-term mating. Men
(vs.~women) had a higher average number of both experiences and mates
when analyses were limited to only heterosexual men and women who
successfully participated in the mating pool (i.e., those with a
non-zero number of short-term mating experiences). Moreover, when men
(vs.~women) had stronger preferences for short-term mating, gay men had
a higher average number of experiences and mates compared to both
lesbian women and heterosexual men. These results suggest that even when
gender differences in mating preferences exist, those in short-term
mating behaviors only occur among particular populations, or when men's
preferences for short-term mating are not constrained by those of women.
Suggestions for future research in human mating psychology and behaviors
were provided.

@@

\emph{Keywords:} agent-based modeling, short-term mating, casual sex,
sexual strategies theory, female choice hypothesis

@@

As a starting point, we took gender differences in mating preferences as
conceptualized by the sexual strategies theory (Buss \& Schmitt, 1993;
Buss \& Schmitt, 2019), which is repeatedly supported by empirical
investigations (e.g., Schmitt, 2003; Walter et al., 2020). Using an
agent-based model, we investigated whether men and women's differential
preferences for short-term mating would result in gender differences in
short-term mating behaviors, and if so, then under what conditions they
would result in differences in the number of short-term mating
experiences and short-term mates.

@@

Gender differences in mating preferences have been studied in the light
of sexual strategies theory (Buss \& Schmitt, 1993; Buss \& Schmitt,
2019). It posits that since the minimum obligatory investment that men
must devote to their offspring (contribution of sperm through one sexual
act) is lower than that of women (gestation, labor, and lactation), men
tend to be relatively more interested in short-term mating compared to
women. This is because the number of offspring women can produce is
limited and cannot be increased by mating with a large number of men,
whereas the reverse is the case for men. As a result, men are predicted
to not only have a greater interest in short-term mating, but also
desire a larger number of short-term mates in a given period and be less
selective with respect to accepting a potential mate (Buss \& Schmitt,
1993).

The hypotheses derived from sexual strategies theory have received
extensive support in empirical studies. For example, men have reported
to be currently seeking short-term mates to a larger extent than women
(Buss \& Schmitt, 1993; Schmitt et al., 2001; Schmitt, 2003), and a
larger proportion of men were in any way seeking short-term mates
(vs.~not seeking) (Schmitt, 2003). Similar results have been found among
both U.S. college students (Buss \& Schmitt, 1993; Schmitt et al., 2001)
as well as in ten world regions in a cross-cultural study (Schmitt,
2003). Similarly, men across the world have also reported to desire more
short-term sexual partners compared to women over different future time
periods (e.g., a month, a year) (Buss \& Schmitt, 1993; McBurney et al.,
2005; Schmitt, 2003). The gender differences have been found regardless
of whether they were estimated using mean (Buss \& Schmitt, 1993;
Schmitt et al., 2001; Schmitt, 2003), median (Schmitt et al., 2001;
Schmitt, 2003), or percentage (Schmitt, 2003) statistics.

As for mating standards, men are less selective in accepting someone as
a potential short-term mate. For example, studies using U.S. college
samples have found that men' acceptable minimum percentile ranks of the
desirability of a potential short-term sexual partner were lower than
those accepted by women, in terms of both the overall desirability and
the desirability of individual traits (e.g., social status,
attractiveness) (Kenrick et al., 1990; Regan, 1998). Another study also
found that when presented with identical descriptions of potential
mates, men on average rated them as more desirable than women did
(Wiederman \& Dubois, 1998).

@@

Since heterosexual sex in most cases involves one man and one woman,
men's relatively high interest in short-term mating can be constrained
by women's preference for long-term relationships (Archer, 2019; Symons,
1979). This is because men's short-term mating preferences -- with the
exception of sexual violence -- can only translate into behaviors when
there are women willing to have sex with them. When a new short-term
mating encounter occurs, both men's and women's total number of
short-term mating experiences and the number of short-term mates
increase by one. Therefore, the total number of short-term mating
experiences and short-term mates must be equal between heterosexual men
and women at the population level. In a closed population with equal sex
ratio, it follows that the average number of experiences and mates
should also be equal between men and women. Following this line of
reasoning, we would expect to observe no gender difference in short-term
mating behaviors among heterosexual individuals even when there were
gender differences in mating preferences.

However, it is important to note that although heterosexual men's total
number of short-term matings would be equal to that of women, this does
not necessarily mean that the total number of men and women who have
ever had short-term mating must be equal. Empirical studies have found
that the proportion of individuals who are sexually inexperienced is
slightly higher among men than women in most age groups above 18 years
old (Ghaznavi et al., 2019; Haydon et al., 2014). Therefore, as a
postulation, there may be a smaller proportion of men (vs.~the
proportion of women) who participate in the mating pool and contribute
to the total number of short-term mating. For example, suppose that
there is a population of 20 individuals with an equal sex ratio.
Heterosexual short-term matings have happened \(n\) times in this
population, but only 3 men and 8 women have had any short-term mating
experiences. Since the total number of short-term mating behaviors is
the same between men and women, we would expect that in a population of
individuals who participate in the mating pool, the average numbers of
short-term mating experiences and mates are higher among men than women
(e.g., in the aforementioned example, the average number of short-term
experiences is \(\frac{n}{3}\) for men and \(\frac{n}{8}\) for women who
are in the mating pool).

As a comparison, in the cases of gay men and lesbian women, men's
preferences are not constrained by those of women's, but only by those
of other men, who, arguably, have a similar high interest in short-term
mating. This would allow for a more direct behavioral expression of
men's mating preferences. The notion that gay men have less restricted
preferences is born out in the proportion of individuals who have
engaged in extradyadic sex (Peplau et al., 2004). A study showed that
the proportions of heterosexual men and women who had engaged in
extradyadic sex were 26\% and 21\%, respectively, while the numbers
among gay men and lesbian women were 82\% and 28\%, respectively (Peplau
et al., 2004). Therefore, we would expect to observe gender differences
in short-term mating behaviors between gay men and lesbian women.
Moreover, we would also expect to find that gay men would engage in more
short-term mating behaviors compared to heterosexual men due to less
constraints on gay men's mating preferences.

@@

We also modeled forming long-term relationships as a background process.
A pair of individuals had a possibility to commit to a long-term
relationship upon meeting each other. This process was modeled at a
behavioral level and the psychological processes behind it was ignored,
although we recognized that gender differences may also exist in
long-term mating strategy (Buss \& Schmitt, 1993; Buss \& Schmitt,
2019).

We modeled these processes among both heterosexual individuals and gay
men and lesbian women to examine whether constraints set by the opposite
sex's preferences would affect short-term mating behaviors. Individuals'
sexual orientation was conceptualized as categorical in terms of
behaviors only in our model. Heterosexual men had short-term mating or
formed a long-term relationship only with women, while gay men only with
other men, and mutatis mutandis for women.

We formulated the following hypotheses: assuming gender differences in
preferences for short-term mating, 1) there would be no gender
differences in short-term mating behaviors between heterosexual men and
women when considering the full population, 2) a smaller proportion of
heterosexual men (vs.~heterosexual women) would have ever engaged in
short-term mating, 3) in a population of heterosexual individuals who
had any short-term mating, men would have a higher average number of
short-term mating behaviors than women, 4) gay men would engage in more
short-term mating behaviors compared to lesbian women, and 5) compared
to heterosexual men.

@@

The model measured two outcomes: (1) the number of short-term mating
experiences of men and women; (2) the number of short-term mates of men
and women. Additionally, we also measured the number of men and women in
the mating pool (i.e., those with a non-zero number of short-term mating
experiences). The average numbers of short-term mating experiences and
short-term mates were calculated by taking the average among the whole
population of men and women (\emph{N\textsubscript{men}} =
\emph{N\textsubscript{women}} = 150) and among those who had ever
engaged in any short-term mating.

@@

At each time step, agents first checked whether they were in a long-term
relationship. If they were in a long-term relationship, they did not
move. If they were not in a long-term relationship, they set their
heading randomly (if they were within the movement range from the
birthplace) or faced the birthplace (if they were out of the movement
range from the birthplace) and moved by a random distance. The random
distance was less than half of the movement range. Then, the agents
decided on whether to engage in short-term mating using the
predetermined likelihood (either 25\% or 40\%).

@@

Then, women randomly selected one man at the same location as their
potential long-term partner. There was a 10\% of chance that a pair of
individuals would form a long-term relationship. After forming a
long-term relationship, both men and women changed to coupled status and
registered each other as their long-term partner.

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=0.8\columnwidth]{figures/fig2_men_vs_women_inpool.pdf}
  \caption{\textbf{Short-term mating behaviors of heterosexual men and women in the mating pool after 1,000 time steps in the model when gender differences existed in mating preferences.} Violin plots summarizing the outcome variables. Plot (A) shows the average number of short-term mating experiences in the mating pool, and plot (B) shows the average number of short-term mates in the mating pool. Central points show mean values and whiskers represent standard errors, but the standard errors are small and overlap to form a single bar. The means were calculated using the population of heterosexual men and women who successfully participated in the mating pool (proportion: \textit{M\textsubscript{men}} = 0.38, \textit{M\textsubscript{women}} = 0.51).}
  \label{fig:men_vs_women_inpool}
\end{figure}

@@

Gay men engaged in short-term mating behaviors more than heterosexual
men did (Figure \ref{fig:hetero_vs_gay}). The average number of
short-term mating experiences (calculated using the full population of
men) was higher among gay men (\emph{M} = 2.63, \emph{SD} = 0.37) than
among heterosexual men (\emph{M} = 1.43, \emph{SD} = 0.24),
\emph{t}(4,998) = --134.87, \emph{p} \textless{} .001, Cohen's \emph{d}
= 3.81, 95\% CI = --1.21 to --1.18. Similarly, the average number of
short-term mates (calculated using the full population of men) was
higher among gay men (\emph{M} = 1.72, \emph{SD} = 0.19) than among
heterosexual men (\emph{M} = 0.80, \emph{SD} = 0.11), \emph{t}(4,998) =
--206.91, \emph{p} \textless{} .001, Cohen's \emph{d} = 5.85, 95\% CI =
--0.92 to --0.91.

\begin{figure}[h]
  \centering
  \includegraphics[width=0.8\columnwidth]{figures/fig3_hetero_vs_gay_men.pdf}
  \caption{\textbf{Short-term mating behaviors of heterosexual men and gay men after 1,000 time steps in the model when gender differences existed in mating preferences.} Violin plots summarizing the outcome variables. Plot (A) shows the average number of short-term mating experiences, and plot (B) shows the average number of short-term mates. Central points show mean values and whiskers represent standard errors, but the standard errors are small and overlap to form a single bar. The means were calculated using the full population of heterosexual men and gay men in the model (\textit{N\textsubscript{hetero}} = \textit{N\textsubscript{gay}} = 150).}
  \label{fig:hetero_vs_gay}
\end{figure}

\hypertarget{experimental-conditions-giving-rise-to-gender-differences-in-short-term-mating-behaviors}{%
\subsection{Experimental conditions giving rise to gender differences in
short-term mating
behaviors}\label{experimental-conditions-giving-rise-to-gender-differences-in-short-term-mating-behaviors}}

@@

\caption{\label{tab:table_1}\textbf{Short-term mating behaviors of heterosexual men and women in the mating pool after 1,000 time steps in the model}}
\centering
\begin{threeparttable}
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{2in}llllllllll}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{5}{c}{\makecell[c]{Average short-term mating\\experiences (in-pool)}} & \multicolumn{5}{c}{\makecell[c]{Average short-term\\mates (in-pool)}} \\
\cmidrule(l{3pt}r{3pt}){2-6} \cmidrule(l{3pt}r{3pt}){7-11}
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Men} & \multicolumn{2}{c}{Women} & \multicolumn{1}{c}{ } & \multicolumn{2}{c}{Men} & \multicolumn{2}{c}{Women} & \multicolumn{1}{c}{ } \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){7-8} \cmidrule(l{3pt}r{3pt}){9-10}
\multicolumn{1}{c}{\em{ }} & \multicolumn{1}{c}{\em{M}} & \multicolumn{1}{c}{\em{SD}} & \multicolumn{1}{c}{\em{M}} & \multicolumn{1}{c}{\em{SD}} & \multicolumn{1}{c}{\em{d}} & \multicolumn{1}{c}{\em{M}} & \multicolumn{1}{c}{\em{SD}} & \multicolumn{1}{c}{\em{M}} & \multicolumn{1}{c}{\em{SD}} & \multicolumn{1}{c}{\em{d}}\\
\midrule
Same likelihood*Same standard & 2.21 & 0.38 & 2.20 & 0.37 & 0.03 & 1.39 & 0.12 & 1.39 & 0.12 & 0.04\\
Same likelihood*Different standard & 2.82 & 0.36 & 2.21 & 0.28 & 1.89 & 1.76 & 0.15 & 1.38 & 0.09 & 3.13\\
Different likelihood*Same standard & 2.80 & 0.48 & 2.78 & 0.47 & 0.03 & 1.57 & 0.14 & 1.56 & 0.14 & 0.05\\
Different likelihood*Different standard & 3.78 & 0.50 & 2.79 & 0.35 & 2.31 & 2.12 & 0.17 & 1.56 & 0.11 & 3.84\\
\bottomrule
\end{tabular}}
\begin{tablenotes}
\small
\item \textit{Note.} The means were calculated with subsamples of men and women who had ever engaged in short-term mating behaviors. See Supplemental Materials for the mean and standard deviation of sample sizes.
\end{tablenotes}
\end{threeparttable}
\end{table}

Among gay men and lesbian women, gender differences in short-term mating
behaviors emerged both when they had different short-term mating
likelihood and when they had different mating standards. Gay men
(vs.~lesbian women) had a higher average number of short-term mating
experiences and short-term mates when they had a higher short-term
mating likelihood (experiences: \emph{M\textsubscript{men}} = 0.80,
\emph{SD\textsubscript{men}} = 0.23, \emph{M\textsubscript{women}} =
0.32, \emph{SD\textsubscript{women}} = 0.10, Cohen's \emph{d} = 2.67;
partners: \emph{M\textsubscript{men}} = 0.52,
\emph{SD\textsubscript{men}} = 0.13, \emph{M\textsubscript{women}} =
0.25, \emph{SD\textsubscript{women}} = 0.07, Cohen's \emph{d} = 2.60),
or when they had lower mating standards (experiences:
\emph{M\textsubscript{men}} = 1.05, \emph{SD\textsubscript{men}} = 0.18,
\emph{M\textsubscript{women}} = 0.32, \emph{SD\textsubscript{women}} =
0.10, Cohen's \emph{d} = 5.10; partners: \emph{M\textsubscript{men}} =
0.82, \emph{SD\textsubscript{men}} = 0.12, \emph{M\textsubscript{women}}
= 0.25, \emph{SD\textsubscript{women}} = 0.07, Cohen's \emph{d} = 5.78).
In comparison, virtually no gender differences in short-term mating
behaviors existed when gay men and lesbian women had the same short-term
mating likelihood and mating standards (Table 2).

@@

The present study used agent-based modeling to investigate whether men
and women's differential preferences for short-term mating would result
in gender differences in short-term mating behaviors, and if so, then
under what conditions they would result in differences in the number of
short-term mating experiences and short-term mates. We formulated five
hypotheses: 1) there would be no gender differences in short-term mating
behaviors between heterosexual men and women when considering the full
population, 2) a smaller proportion of heterosexual men
(vs.~heterosexual women) would have ever engaged in short-term mating,
3) in a population of heterosexual individuals who had any short-term
mating, men would have a higher average number of short-term mating
behaviors than women, 4) gay men would engage in more short-term mating
behaviors compared to lesbian women, and 5) compared to heterosexual
men. The results from 1,000 time steps in a model simulating men and
women's mating behaviors provided strong evidence in favor of our
hypotheses. First of all, we found no gender differences in short-term
mating behaviors between heterosexual men and women when the full
population was considered. However, compared to heterosexual women,
there was a smaller proportion of heterosexual men participating in the
mating pool, but they had higher average numbers of short-term mating
experiences and short-term mates. Secondly, gay men had higher average
numbers of short-term mating experiences and short-term mate compared to
both lesbian women and heterosexual men.

As expected, when all individuals in the population were considered,
heterosexual men and women did not differ in short-term mating behaviors
despite their differential preferences for short-term mating. This is
not surprising since the sex ratio was 1:1 in our model. Heterosexual
men and women had an equal total number of both short-term mating
experiences and short-term mates. It therefore follows that the average
number of short-term mating experiences and short-term mates must be
equal between men and women as well. However, we did find that when we
only looked at individuals in the mating pool (from which more men than
women were excluded), men engaged in more short-term mating behaviors
compared to women. This was because there were more women than men in
the mating pool, resulting in lower averages among heterosexual women
despite the equal numbers of experiences and mates in total.
